New insights into the Weddell Sea ecosystem applying a quantitative network approach

### 2022-12-30 Manuscript/Manuscript.Rmd
Submitted version

correspongdingauthors: Tomás I. Marina (tomasimarina@gmail.com) and Leonardo A. Saravia (arysar@gmail.com)
# If authors contributed equally, please mark the respective author names with an asterisk '*' and add a further affiliation: 'These authors contributed equally to this work.'", see template.
affiliation:
  - code: 1
    address: Centro Austral de Investigaciones Científicas (CADIC-CONICET), Ushuaia, Argentina
  - code: 2
    address: Spatial Foodweb Ecology Group, Department of Agricultural Sciences, University of Helsinki, Finland
  - code: "*"
    address: These authors contributed equally to this work.
abstract: |
  Network approaches can shed light on the structure and stability of complex marine communities. In recent years, such approaches have been successfully applied to study polar ecosystems, improving our knowledge on how they might respond to ongoing environmental changes. The Weddell Sea is one of the most studied marine ecosystems outside the Antarctic Peninsula in the Southern Ocean. Yet, few studies consider the known complexity of the Weddell Sea food web, which in its current form comprises 490 species and 16041 predator-prey interactions. Here we analysed the Weddell Sea food web, focusing on the species and trophic interactions that underpin ecosystem structure and stability. We estimated the strength for each interaction in the food web, characterised species position in the food web using unweighted and weighted food web properties, and analysed species’ roles with respect to the stability of the food web. We found that the distribution of the interaction strength at the food web level is asymmetric, with many weak interactions and few strong ones. We detected a positive relationship between species mean interaction strength and two unweighted properties (i.e., trophic level and the total number of interactions). We also found that only a few species possess key positions in terms of food web stability. These species are characterised by high mean interaction strength, mid to high trophic level, relatively high number of interactions, and mid to low trophic similarity. In this study, we integrated unweighted and weighted food web information, enabling a more complete assessment of the ecosystem structure and function of the Weddell Sea food web. Our results provide new insights, which are important for the development of effective policies and management strategies, particularly given the ongoing initiative to implement a Marine Protected Area (MPA) in the Weddell Sea.

@@

acknowledgements: |
  Thanks to the rticles contributors!
# appendix: |
#   \section{Figures and tables in appendices}

@@

# Introduction

<!-- Relevance of studying food webs -->
Food web analysis constitutes an important framework for understanding ecological community structure and for conserving biodiversity through ecosystem management [@Thompson2012]. Although topological food web analysis, which considers only the presence and absence of predator-prey interactions, provides important insights into the structure and functioning of ecological communities [e.g. @Pascual2006; @Kortsch2015; @Marina2018; @Cordone2020; @Rodriguez2022], more information on the nature of the trophic interactions is needed to effectively charaterise ecosystem dynamics and stability [e.g. @Kortsch2021; @Pecuchet2022]. This is a fundamental step for providing assemessments on ecosystem vulnerability to environmental pressures and for prioritising management actions. In this regard, quantifying the strength of trohic interactions and species' roles within the network are of paramount importance [@Carrara2015; @Allesina2015; @Nilsson2016; @Cirtwill2018a].

<!-- Paragraph on unweighted and weighted networks  -->
Estimating interaction strength in food webs allows differentiating the importance of species interactions. On the contrary, unweighetd food web representations give equal importantance to all interactions well-knowing that some species interactions are stronger than others and hence play a diffrent role for ecosystem functioning and stability. Both empirical and theoretical studies show that interactions strength distributions in food webs are assymetric [@Paine1992; @McCann1998; @Emmerson2004; @Wootton2005; @Kortsch2021], containing a few strong and many weak interactions. This asymmetric patterning of weak and strong links is crucial to food web stability [@Paine1992; @McCann1998; @Neutel2002]. In a recent paper on an aquatic food web it was further highlighted that temporal changes in ecosytem functioning could only be predicted using weighted food web structure [@Kortsch2021]. Hence, in order to assess the stability and functioning of a food web, it is important to first determine the quantitative structure of the trophic network.

<!-- LAS I think that the methodology of estimating the IS is in the heart of the paper so we need to say something about that, this is why I added the following paragraph-->
Several methodologies have been applied to estimate interaction strength in food webs, where the quantity and quality of the data mostly determines which approach is the most convenient [@Berlow2004]. Approaches include experimental methods combined with dynamic modelling [@Emmerson2004; @Carrara2015], measurements of species abundances through time [@Fahimipour2014; @Chang2021], and estimation of metabolic rates and biomass of all species in the community [@Neutel2014]. However, these types of methods require large experimental set-ups and parameterisations restricting the analyses to smaller networks (e.g., approximatley 10 species or less). Other methods based on allometric scaling relationships and biomass information [@Kortsch2021; @Gauzens2019] can be applied to larger networks with less data requirements, but this comes at the expense of precision in the predictions. For even larger food webs composed of nearly 1000 species and more than 10000 interactions, only methods with even less data requirements are feasible. One of these methods, proposed by @Pawar2012, combines data on consumer and resource body masses and consumer search space (interaction dimensionality) to obtain interaction strength estimates for each pairwise predator-prey interaction. An advantage of this method is that it can be applied without information on species biomass.

<!-- Insights of network science to polar ecosystems -->
Using a a network approach, different types (e.g., terrestial, lake, marine) of food webs from various geographic locations are studied worldwide, including marine polar food webs [@Carscallen2012; @Santana2013; @Kortsch2019; @Pecuchet2022]. Some of the studies from the Arctic show how food web properties (e.g., connectance) are constrained by environmental factors such as sea ice cover and seawater temperature [@Kortsch2019; @Pecuchet2022]. In the Southern Ocean, important insights have been gained into mechanisms of energy flow, the relative importance of individual species and their traits, and the influence of environmental variables (e.g., sea-ice) on the structure of local food webs [@Cordone2020; @Rossi2019]. For instance, in Potter Cove (West Antarctic Peninsula) the substratum type (i.e. hard/soft or rocks/sediments) plays a significant role in the structure and stability of the food web. In Terra Nova Bay (Ross Sea), the architecture of biodiversity was reshaped by the pulsed input of sympagic food sources following sea-ice break up, with food web simplification, decreased intraguild predation, potential disturbance propagation and increased vulnerability to biodiversity loss [@Rossi2019].

<!-- Climate change in Weddell Sea -->
The Weddell Sea is expected to be one of the last regions of the Southern Ocean to experience the consequences of climate change due to its extensive ice cover and ocean currents [@Teschke2021] resulting in less sea surface warming compared to other areas of the Southern Ocean. This Sea plays an important role in driving global thermohaline circulation and ventilating the global abyssal ocean because it generates a considerable part of the Antarctic Bottom Water [@Fahrbach2009]. Because of these environmental characteristics, the Weddell Sea may serve as a refuge for Antarctic species which depend on sea ice (e.g. krill, emperor penguin, Weddell seal) or have low heat tolerance (e.g., most notothenioid fishes) due to their adaptations to freezing tempeartures [@Griffiths2017]. While essential large-scale hydrodynamic relationships are relatively well-known for this region [@deSteur2019], information on the current distribution, abundance and sensitivity to climate change is only partially known for a few species (e.g., emperor penguin) [@Houstin2022].

<!-- Food web studies about Weddell Sea / Key species or species role -->
The network complexity of the Weddell Sea food web is high comprising 488 species and 16200 predator-prey interactions [@Jacob2011]. In an attempt to better understand species roles related to food web stability, Jacob et al. (2011) performed secondary extinction experiments and found that the removal of small to medium-sized, and not large, organisms caused a cascade of secondary extinctions. This findings highlighted the relative importance of predators, rather than prey, for the architecture, functioning and stability of the Weddell Sea food web, which coincides with findings from recent meta-analyses in natural complex food webs [@Brose2019; @Perkins2022]. Other investigations considered this food web in a meta-analysis context showing that high predator-prey body-mass ratios are found for predator groups with specific trait combinations, including small vertebrates and large swimming or flying predators [@Brose2019]. These trait combinations generate weak interactions that stabilize communities against perturbations maintaining ecosystem functioning.

<!-- Objectives -->
In this study, we aim to go beyond a purely topological (presence/absence) assessment of who eats whom in the Weddell Sea ecosystem by providing a quantitative analysis of the trophic interaction network. We aim to analyse the species' role for the structure and stability of the food web. To achieve this, we: 1) estimated the strength for each interaction in the Weddell Sea food web, 2) characterised species' role considering both weighted and unweighted properties, and 3) analysed the species’ role related to the stability of the food web. This is the first time that interaction strengths were estimated for all pairwise trophic interactions at the species level (except for a few) for the Weddell sea food web.

@@

We fitted six candidate models (Exponential, Gamma, log-Normal, Normal, Power-law and Uniform) the interaction strength distribution using maximum likelihood [@McCallum2008], and selected the best fitting model by computing the Akaike Information Criterion AIC [@Burnham2002].

@@

To assess species roles as a function of the weighted food web, we focused on mean interaction strength defined as the average strength of all interactions for a given species. Further we calculated three unweighted species properties: a) species degree, i.e., the sum of in- and out-going interactions ; b) trophic level ; and c) trophic similarity, i.e., the trophic overlap based on shared and unique resources and consumers. These metrics were chosen to assess a species role based on the unweighted network. The species degree has often been equated with species importance to the structure and functioning within a food web, i.e. perturbations to high-degree species may therefore have more significant effects on the food web robustness to perturbations than low-degree species [@Dunne2002a; references in @Cirtwill2018a]. The trophic level offers information about how important a species is to its biotic community, i.e., top predators and primary producers are expected to have particularly large effects on the rest of their communities through top-down and bottom-up control, respectively [references in @Cirtwill2018a]. Trophic similarity is an index of trophic overlap considering the set of prey and predators for a pair of species; it measures one of the most important aspects of species’ niches, the trophic niche, and functional aspects of biodiversity [@Martinez1991; @Williams2000].

@@

To study the relationship between species mean interaction strength (weighted property) and the unweighted species properties, we performed linear regression analyses between the log mean interaction strength and each of the aforementioned unweighted properties. Thus, we considered the interaction strength as the dependent variable and the given unweighted property as the independent variable, and obtained the coefficients (slope and intercept) for the linear model. Models were fitted using the least squares approach. We also explored the mean interaction strength distribution with the species habitat.

@@

To identify the species with the highest effect on food web stability, we plotted the results of each species' extinction and its effect on food web stability, considering weighted (interaction strength) and unweighted properties, and species habitat.

@@

The statistical distribution that best fitted the empirical interaction strength distribution of the Weddell Sea food web was a ‘gamma’ due to the high proportion of weak interactions and the existence of a few strong interactions (Figure 3, Table S3). <!-- Gamma parameters ?? [Tomás] I added them in Sup Mat -->

@@

We found that the species’ mean interaction strength (weighted property) shows different relationships with the unweighted properties analysed (Figure 4A-D). In this regard, there is a positive relationship between interaction strength and trophic level, i.e., the higher the trophic level of the species, the higher its mean interaction strength. We also found a significant but less evident positive relationship with species degree. Contrary, there was no significant relationship between mean interaction strength and trophic similarity. Considering species habitat affiliation, the “Benthopelagic” and “Pelagic” categories contained the two species with the highest mean interaction strength, the killer whale Orcinus orca and the colossal squid Mesonychoteuthis hamiltoni, respectively. However, the majority of the species with relatively higher interaction strength belonged to the “Demersal” and “Land-based” habitats groups. Species inhabiting the benthic realm showed the lowest mean interaction strength (Figure 4D).

@@

Our extinction analyses showed that the majority of species had no significant impact on food web stability after being removed (Figure 5). Most of the species (black points in figure 5) did not change the stability of the network considerably after being removed, except for a few species. Only 15 out of 490 species (3.06%) gave rise to significant changes in the food web’s stability after their removal (Table 2). Most of these species had a positive impact on food web stability, i.e., network stability increased after their removal. Only two species significantly decreased network stability after being removed, the demersal fish Pagetopsis macropterus and the benthopelagic amphipod Maxilliphimedia longipes.

After exploring the stability difference against the species properties (Figure 5), we found that the species that generated a significant impact on the stability of the food web were characterised by: 1) high mean interaction strength; 2) mid to high trophic levels (TL > 3.2); 3) relatively high number of interactions (Degree > 25); and 4) mid to low trophic similarity (TS < 0.16). Habitat wise, species with a significant impact on the stability were present in all habitats, except for the benthic realm. Table 2 shows the results for the species with highest impact on the food web stability.

@@

Our analyses show that the distribution of species interaction strength (IS) at the network level is asymmetric, i.e., the Weddell Sea food web contains many weak interactions and only a few strong ones. This finding is consistent with many previous theoretical and empirical studies [e.g. @McCann1998; @Neutel2002; @Emmerson2004; @Wootton2005; @Kortsch2021]. The asymmetric distribution of IS in food webs has been interpreted as an explanation for the persistence of complex communities in nature [@Bascompte2005; @Allesina2015; @Nilsson2016]. Here we show that this pattern is also prevalent in the Weddell Sea, one of the most complex food webs to date, comprising 490 species and 16041 predator-prey interactions. This finding is in someway validating the method we used, validation of allometric methods of IS estimation not including interaction dimensionality has been performed for microcosmos of relatively few species [@Jonsson2018].

@@

Our study goes beyond the current understanding of how species influence ecosystem structure and stability in the Weddell Sea in particular, and in most polar regions in general [@Murphy2016; @McCormack2021]. In the same analysis we integrated information on weighted (IS) and unweighted species properties, enabling a more complete assessment of the species’ role realted food web structure and stability, and allowing us to identify species and their characteristics which can have a destabilising or stabilising effect on the food web.

@@

<!-- LAS Figure 2 is confusing, on the left side the title is "Weighted" but the network seems unweighted, I would delete the small predator-prey on the left, and explain that arrow are the flux of energy thus pointin to the predator, and that IS is represented by the width of the arrow. -->

@@

<!-- EXTRA FROM INTRO -->
<!-- Interaction strength -->
<!-- Several methodologies have been applied to estimate the interaction strength in food webs, where the quantity and quality of data mostly determine which approach is more convenient [@Berlow2004]. Experimental methods combined with dynamic modelling [@Emmerson2004; @Carrara2015], measurements of species abundances trough time [@Fahimipour2014; @Chang2021], and/or estimation of metabolic rates and biomass of all species in the community [@Neutel2014]; all these methods require a large experimental and/or sampling effort that restrict them to simple networks (circa 10 species). Other methods based on predator-prey body mass allometric relationships [@Berlow2009; @Pawar2012] could be applied to larger networks due to less data requirements, although they might have less power with increasingly complex networks unless more traits are included in the estimations [@Jonsson2018]. For food webs composed by nearly 1000 species and about 20000 interactions only those methods with less data requirements are feasible. On of these is the one proposed by @Pawar2012, which allows to obtain interaction strength estimations by combining data on body mass (of consumer -predator- and resource -prey-) and interaction dimensionality (classified as 2 or 3 dimensions). -->

<!-- The Weddell Sea represents the southerly part of the Atlantic Sector of the Southern Ocean. About one quarter of the Weddell Sea’s entire marine area lies over the continental shelf, stretching along the eastern contour of the Antarctic Peninsula and the Antarctic continent up to 20◦E. The ecosystem of the Weddell Sea is one of the most pristine and relatively unexplored hotspots of marine biodiversity in Antarctica [@Teschke2021]. It contains several areas of ecological significance for birds and marine mammals, such as important breeding and foraging areas [@Hindell2020]. Furthermore, the sea floor is home to large and unique forest-like sponge and filter-feeder assemblages, which are particularly numerous on the continental shelf in the eastern part of the Weddell Sea [@Pineda-Metz2019]. These species assemblages create a structurally complex and diverse habitat for thousands of species [@Barthel1992; @Brey1994; Brandt2007]. -->

>>>>

correspongdingauthors: Tomás I. Marina (tomasimarina@gmail.com) and Leonardo A. Saravia (arysar@gmail.com)

affiliation:
  - code: 1
    address: Centro Austral de Investigaciones Científicas (CADIC-CONICET), Ushuaia, Argentina
  - code: 2
    address: Spatial Foodweb Ecology Group, Department of Agricultural Sciences, University of Helsinki, Finland
  - code: "*"
    address: These authors contributed equally to this work.
abstract: |
  Network approaches can shed light on the structure and stability of complex marine communities. In recent years, such approaches have been successfully applied to study polar ecosystems, improving our knowledge on how they might respond to ongoing environmental changes. The Weddell Sea is one of the most studied marine ecosystems outside the Antarctic Peninsula in the Southern Ocean. Yet, few studies consider the known complexity of the Weddell Sea food web, which in its current form comprises 490 species and 16041 predator-prey interactions. Here we analysed the Weddell Sea food web, focusing on the species and trophic interactions that underpin ecosystem structure and stability. We estimated the strength for each interaction in the food web, characterised species position in the food web using unweighted and weighted food web properties, and analysed species’ roles with respect to the stability of the food web. We found that the distribution of the interaction strength (IS) at the food web level is asymmetric, with many weak interactions and few strong ones. We detected a positive relationship between species mean IS and two unweighted properties (i.e., trophic level and the total number of interactions). We also found that only a few species possess key positions in terms of food web stability. These species are characterised by high mean IS, mid to high trophic level, relatively high number of interactions, and mid to low trophic similarity. In this study, we integrated unweighted and weighted food web information, enabling a more complete assessment of the ecosystem structure and function of the Weddell Sea food web. Our results provide new insights, which are important for the development of effective policies and management strategies, particularly given the ongoing initiative to implement a Marine Protected Area (MPA) in the Weddell Sea.

@@

acknowledgements: |
  We thank all persons that dedicated time to build and curate the dataset GATEWAy that enable us to carry out this work. We also thank the organizers of the different virtual workshops of the Weddell Sea and Droning Maud Land (WSDML) Regional Working Group for giving us the possibility to participate and contribute in the Special Issue.

@@

\introduction[Introduction]

Food web analysis constitutes an important framework for understanding ecological community structure and for conserving biodiversity through ecosystem management [@Thompson2012]. Although topological food web analysis, which considers only the presence and absence of predator-prey interactions, provides important insights into the structure and functioning of ecological communities [e.g. @Pascual2006; @Kortsch2015; @Marina2018; @Cordone2020; @Rodriguez2022], more information on the nature of the trophic interactions is needed to effectively charaterise ecosystem dynamics and stability [e.g. @Kortsch2021; @Pecuchet2022]. This is a fundamental step for providing assemessments on ecosystem vulnerability to environmental pressures and for prioritising management actions. In this regard, quantifying the strength of trohic interactions and species' roles within the network are of paramount importance [@Carrara2015; @Allesina2015; @Nilsson2016; @Cirtwill2018a].

Estimating interaction strength (IS) in food webs allows differentiating the importance of species interactions. On the contrary, unweighetd food web representations give equal importantance to all interactions well-knowing that some species interactions are stronger than others and hence play a diffrent role for ecosystem functioning and stability. Both empirical and theoretical studies show that interactions strength distributions in food webs are assymetric [@Paine1992; @McCann1998; @Emmerson2004; @Wootton2005; @Kortsch2021], containing a few strong and many weak interactions. This asymmetric patterning of weak and strong links is crucial to food web stability [@Paine1992; @McCann1998; @Neutel2002]. In a recent paper on an aquatic food web it was further highlighted that temporal changes in ecosytem functioning could only be predicted using weighted food web structure [@Kortsch2021]. Hence, in order to assess the stability and functioning of a food web, it is important to first determine the quantitative structure of the trophic network.

Several methodologies have been applied to estimate IS in food webs, where the quantity and quality of the data mostly determines which approach is the most convenient [@Berlow2004]. Approaches include experimental methods combined with dynamic modelling [@Emmerson2004; @Carrara2015], measurements of species abundances through time [@Fahimipour2014; @Chang2021], and estimation of metabolic rates and biomass of all species in the community [@Neutel2014]. However, these types of methods require large experimental set-ups and parameterisations restricting the analyses to smaller networks (e.g., approximatley 10 species or less). Other methods based on allometric scaling relationships and biomass information [@Kortsch2021; @Gauzens2019] can be applied to larger networks with less data requirements, but this comes at the expense of precision in the predictions. For even larger food webs composed of nearly 1000 species and more than 10000 interactions, only methods with even less data requirements are feasible. One of these methods, proposed by @Pawar2012, combines data on consumer and resource body masses and consumer search space (interaction dimensionality) to obtain IS estimates for each pairwise predator-prey interaction. An advantage of this method is that it can be applied without information on species biomass.

Using a a network approach, different types (e.g., terrestial, lake, marine) of food webs from various geographic locations are studied worldwide, including marine polar food webs [@Carscallen2012; @Santana2013; @Kortsch2019; @Pecuchet2022]. Some of the studies from the Arctic show how food web properties (e.g., connectance) are constrained by environmental factors such as sea ice cover and seawater temperature [@Kortsch2019; @Pecuchet2022]. In the Southern Ocean, important insights have been gained into mechanisms of energy flow, the relative importance of individual species and their traits, and the influence of environmental variables (e.g., sea-ice) on the structure of local food webs [@Cordone2020; @Rossi2019]. For instance, in Potter Cove (West Antarctic Peninsula) the substratum type (i.e. hard/soft or rocks/sediments) plays a significant role in the structure and stability of the food web. In Terra Nova Bay (Ross Sea), the architecture of biodiversity was reshaped by the pulsed input of sympagic food sources following sea-ice break up, with food web simplification, decreased intraguild predation, potential disturbance propagation and increased vulnerability to biodiversity loss [@Rossi2019].

The Weddell Sea is expected to be one of the last regions of the Southern Ocean to experience the consequences of climate change due to its extensive ice cover and ocean currents [@Teschke2021] resulting in less sea surface warming compared to other areas of the Southern Ocean. This Sea plays an important role in driving global thermohaline circulation and ventilating the global abyssal ocean because it generates a considerable part of the Antarctic Bottom Water [@Fahrbach2009]. Because of these environmental characteristics, the Weddell Sea may serve as a refuge for Antarctic species which depend on sea ice (e.g. krill, emperor penguin, Weddell seal) or have low heat tolerance (e.g., most notothenioid fishes) due to their adaptations to freezing tempeartures [@Griffiths2017]. While essential large-scale hydrodynamic relationships are relatively well-known for this region [@deSteur2019], information on the current distribution, abundance and sensitivity to climate change is only partially known for a few species (e.g., emperor penguin) [@Houstin2022].

The network complexity of the Weddell Sea food web is high comprising 488 species and 16200 predator-prey interactions [@Jacob2011]. In an attempt to better understand species roles related to food web stability, Jacob et al. (2011) performed secondary extinction experiments and found that the removal of small to medium-sized, and not large, organisms caused a cascade of secondary extinctions. This findings highlighted the relative importance of predators, rather than prey, for the architecture, functioning and stability of the Weddell Sea food web, which coincides with findings from recent meta-analyses in natural complex food webs [@Brose2019; @Perkins2022]. Other investigations considered this food web in a meta-analysis context showing that high predator-prey body-mass ratios are found for predator groups with specific trait combinations, including small vertebrates and large swimming or flying predators [@Brose2019]. These trait combinations generate weak interactions that stabilize communities against perturbations maintaining ecosystem functioning.

In this study, we aim to go beyond a purely topological (presence/absence) assessment of who eats whom in the Weddell Sea ecosystem by providing a quantitative analysis of the trophic interaction network. We aim to analyse the species' role for the structure and stability of the food web. To achieve this, we: 1) estimated the strength for each interaction in the Weddell Sea food web, 2) characterised species' role considering both weighted and unweighted properties, and 3) analysed the species’ role related to the stability of the food web. This is the first time that interaction strengths were estimated for all pairwise trophic interactions at the species level (except for a few) for the Weddell sea food web.

@@

We fitted six candidate models (Exponential, Gamma, log-Normal, Normal, Power-law and Uniform) the IS distribution using maximum likelihood [@McCallum2008], and selected the best fitting model by computing the Akaike Information Criterion AIC [@Burnham2002].

@@

To assess species roles as a function of the weighted food web, we focused on mean IS defined as the average strength of all interactions for a given species. Further we calculated three unweighted species properties: a) species degree, i.e., the sum of in- and out-going interactions ; b) trophic level ; and c) trophic similarity, i.e., the trophic overlap based on shared and unique resources and consumers. These metrics were chosen to assess a species role based on the unweighted network. The species degree has often been equated with species importance to the structure and functioning within a food web, i.e. perturbations to high-degree species may therefore have more significant effects on the food web robustness to perturbations than low-degree species [@Dunne2002a; references in @Cirtwill2018a]. The trophic level offers information about how important a species is to its biotic community, i.e., top predators and primary producers are expected to have particularly large effects on the rest of their communities through top-down and bottom-up control, respectively [references in @Cirtwill2018a]. Trophic similarity is an index of trophic overlap considering the set of prey and predators for a pair of species; it measures one of the most important aspects of species’ niches, the trophic niche, and functional aspects of biodiversity [@Martinez1991; @Williams2000].

@@

To study the relationship between species mean IS (weighted property) and the unweighted species properties, we performed linear regression analyses between the log mean IS and each of the aforementioned unweighted properties. Thus, we considered the IS as the dependent variable and the given unweighted property as the independent variable, and obtained the coefficients (slope and intercept) for the linear model. Models were fitted using the least squares approach. We also explored the mean IS distribution with the species habitat.

@@

To identify the species with the highest effect on food web stability and their characteristics, we plotted the results of each species' extinction and its effect on food web stability (the stability difference) against weighted (interaction strength), and the three unweighted properties (trophic level, degree, and trophic similarity), and  species habitat affiliation.

@@

The statistical distribution that best fitted the empirical interaction strength distribution of the Weddell Sea food web was a ‘gamma’ due to the high proportion of weak interactions and the existence of a few strong interactions (Figure 3, Table S3).

@@

We found that the species’ mean IS (weighted property) shows different relationships with the unweighted properties analysed (Figure 4A-D). In this regard, there is a positive relationship between IS and trophic level, i.e., the higher the trophic level of the species, the higher its mean IS. We also found a significant but less evident positive relationship with species degree. Contrary, there was no significant relationship between mean IS and trophic similarity. Considering species habitat affiliation, the “Benthopelagic” and “Pelagic” categories contained the two species with the highest mean IS, the killer whale Orcinus orca and the colossal squid Mesonychoteuthis hamiltoni, respectively. However, the majority of the species with relatively higher IS belonged to the “Demersal” and “Land-based” habitats groups. Species inhabiting the benthic realm showed the lowest mean IS (Figure 4D).

@@

Our extinction analyses showed that the majority of species had no significant impact on food web stability after being removed (Figure 5). Most of the species (black points in figure 5) did not change the stability of the network considerably after being removed, except for a few species (red points in figure 5). Only 15 out of 490 species (3.06%) gave rise to significant changes in the food web’s stability after their removal (Table 2). Network stability increased after the removal of most of these species, i.e., these species have a negative effect on stability. Only two species significantly decreased network stability after being removed, the demersal fish Pagetopsis macropterus and the benthopelagic amphipod Maxilliphimedia longipes.

After exploring the stability difference against the species properties (Figure 5), we found that the species that generated a significant impact on the stability of the food web were characterised by: 1) high mean IS; 2) mid to high trophic levels (TL > 3.2); 3) relatively high number of interactions (Degree > 25); and 4) mid to low trophic similarity (TS < 0.16). Habitat wise, species with a significant impact on the stability were present in all habitats, except for the benthic realm. Table 2 shows the results for the species with highest impact on the food web stability.

@@

Our analyses show that the distribution of species IS at the network level is asymmetric, i.e., the Weddell Sea food web contains many weak interactions and only a few strong ones. This finding is consistent with many previous theoretical and empirical studies [e.g. @McCann1998; @Neutel2002; @Emmerson2004; @Wootton2005; @Kortsch2021]. The asymmetric distribution of IS in food webs has been interpreted as an explanation for the persistence of complex communities in nature [@Bascompte2005; @Allesina2015; @Nilsson2016]. Here we show that this pattern is also prevalent in the Weddell Sea, one of the most complex food webs to date, comprising 490 species and 16041 predator-prey interactions. This finding is in someway validating the method we used, validation of allometric methods of IS estimation not including interaction dimensionality has been performed for microcosmos of relatively few species [@Jonsson2018].

@@

Our study goes beyond the current understanding of how species influence ecosystem structure and stability in the Weddell Sea, in particular, and in most polar regions in general [@Murphy2016; @McCormack2021]. In the same analysis we integrated information on weighted (IS) and unweighted species properties, enabling a more complete assessment of species’ role with respect to food web structure and stability. Futher, tha analyses allowed us to identify species and their characteristics which can have a destabilising or stabilising effect on the food web.

@@

@@
